---
title: "Store Location Optimization"
author: "Om Prakash Bhandari"
date: "20/08/2024"
output: workflowr::wflow_html
---

## Geomarketing

Geomarketing is a marketing strategy that involves the integration of geographic information into marketing efforts to make data-driven decisions. It leverages geographic data, such as location, demographic information, and regional behavior patterns, to optimize marketing strategies, target specific audiences, and enhance decision-making processes.

### 1. Load required libraries

We use sf for spatial data manipulation, dplyr for data wrangling, caret for model
training and evaluation, randomForest for the predictive model, and geosphere for
geographic distance calculations.

```{r libraries, message=FALSE, warning=FALSE, include=TRUE}
library(sf)
library(dplyr)
library(caret)
library(randomForest)
library(geosphere)
```


### 2. Data Loading and Preparation

We load the sales, demographic, and geographic data, and merge them into a single dataset.

```{r dataPrep, message=FALSE, warning=FALSE, include=TRUE}
sales_data <-
  read.csv("data/sales_data.csv")
demographic_data <-
  read.csv("data/demographic_data.csv")
geodata <-
  read_sf("data/german_cities.gdb")

# Merge sales and demographic data
data <-
  merge(sales_data, demographic_data, by = "city_name")

# Select the desired columns and rename them
data <- data %>%
  select(city_name,
    latitude = latitude.x, longitude = longitude.x, store_id,
    sale_date, sales_amount, population_density, income_level, average_age
  )

# Convert to spatial data frame
data_sf <-
  st_as_sf(data, coords = c("longitude", "latitude"), crs = 4326)
```


### 3. Feature Engineering 

We calculate the distance to existing stores for each data point.

```{r featureEngineering, message=FALSE, warning=FALSE, include=TRUE}
dist_to_existing_stores <-
  function(lat, lon, existing_locations) {
    distances <- distHaversine(cbind(lon, lat), existing_locations)
    return(mean(distances))
  }

# Add latitude and longitude as separate columns to the sf object
coords <- st_coordinates(data_sf)
data_sf$longitude <- coords[, "X"]
data_sf$latitude <- coords[, "Y"]

existing_locations <- st_coordinates(data_sf)
data_sf <- data_sf %>%
  rowwise() %>%
  mutate(
    distance_to_existing_stores =
      dist_to_existing_stores(latitude, longitude, existing_locations)
  )
```


### 4. Model Building

We train a Random Forest model to predict sales based on the features.

```{r model, message=FALSE, warning=FALSE, include=TRUE}
# Prepare data for modeling
features <- data_sf %>%
  select(distance_to_existing_stores, population_density, income_level)

sales <- data_sf$sales_amount

# split data into training and test sets
set.seed(42)
train_index <- createDataPartition(sales, p = 0.7, list = FALSE)
train_data <- features[train_index, ]
train_sales <- sales[train_index]

test_data <- features[-train_index, ]
test_sales <- sales[-train_index]

# train random forest model
# remove non numeric column from train data

train_data_nosf <-
  st_drop_geometry(train_data)


model <- randomForest(
  x = train_data_nosf,
  y = train_sales,
  ntree = 100,
  importance = TRUE
)
```

### 5. Optimization

We create a function to find the optimal location by predicting sales for potential new locations.

```{r optimization, message=FALSE, warning=FALSE, include=TRUE}
# example potential locations
# Hanover, Nuremberg, Dresden
potential_locations <- data.frame(
  latitude = c(52.3759, 49.4543, 51.0504),
  longitude = c(9.7320, 11.0746, 13.7373),
  population_density = c(4125, 4325, 3925),
  income_level = c(46000, 48000, 44000)
)

potential_locations_new <- potential_locations %>%
  rowwise() %>%
  mutate(distance_to_existing_stores = dist_to_existing_stores(
    latitude, longitude,
    existing_locations
  ))

predictions <- predict(model, newdata = as.data.frame(potential_locations_new))
optimal_index <- which.max(predictions)

optimal_location <- potential_locations_new[optimal_index, ]
print(optimal_location)
```

### 6. Validation

We evaluate the model performance using RMSE.

```{r validation, message=FALSE, warning=FALSE, include=TRUE}
# predict on data set
predictions <- predict(model, newdata = as.data.frame(test_data))

# calculate accuracy metrics
mse <- mean((predictions - test_sales)^2)
rmse <- sqrt(mse)
cat("Root Mean Squared Error:", rmse)
```

### 7. Predicted sales

We print the optimal location along with the predicted sales.

```{r predict, message=FALSE, warning=FALSE, include=TRUE}
optimal_location$predicted_sales <-
  predict(model, newdata = as.data.frame(optimal_location))

print(optimal_location)
```

This example above provides a comprehensive approach to Geomarketing analysis using R.








